Leadership across cultures and contexts: an exploratory analysis

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Re worked discussion on leader qualities cluster.

Several of these traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*. The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return. As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p. 85]; and among the Highland Scots, @es10-003 [p. 147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions. @nm09-056 [p. 183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:

@@

The subcluster of social cost variables, which included *Physically formidable*, *Aggressiveness*, and *Feared*, align fairly closely with the *Dominance* theory of status and leadership [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013]. Interestingly, this cluster also contained *Supernatural*. There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.

>>>>

The three most frequent leader quality variables were found in the *Individual qualities* cluster, which had two smaller and strongly supported subclusters: *Social, reproductive, and material success*, which included traits indicative of the benefits of successful leadership, and a more general *Competencies* cluster which included qualities that are not necesssarily prosocial. Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme. Several of these measures of success often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The *Competencies* subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return. Interestingly, *Feared* and *Supernatural* clustered more closely with these prestige traits and not among the dominance traits. There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership. Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p. 85]; and among the Highland Scots, @es10-003 [p. 147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions. @nm09-056 [p. 183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:

@@
